### 2018-05-02 Paper/bg.tex
work in progress ...

\subsection{Derivative-based global sensitivity analysis}  
Let $\mathcal{Y}$ be a mathematical model that is a function of $\Np$ uncertain 
parameters, $\theta_1, \theta_2, \ldots, \theta_\Np$. The goal of sensitivity analysis
is measuring the influence of each component of the parameter vector 
$\bm{\theta}$ on the model output. 
\alennote{We consider the case of statistically independent parameters, right? 
Should we make this precise?}
Derivative-based global sensitivity analysis (DGSA) is performed by 
computing derivative based global sensitivity measures (DGSMs)~\cite{Sobol:2009} 
for each uncertain parameter in the model. 
Specifically, we consider the following DGSMs, 
\be
\mu_i = 
\mathbb{E}\left[\left(\frac{\partial \mathcal{Y}(\bm{\bm{\theta}})}{\partial \theta_i}\right)^{2}\right], \quad i = 1, \ldots, \Np.
\label{eq:mu}
\ee
Here $\mathbb{E}$ denotes expectation over the uncertain parameters.
Notice that this formulation assumes that the function $\mathcal{Y}$ is differentiable
with respect to $\theta_i$, $i = 1, \ldots, \Np$, almost surely.

@@

\be
\frac{\partial \mathcal{Y}(\bm{\theta}^{\ast})}{\partial \theta_i} =
 \lim_{\Delta\theta_i^{*}\to 0}
\frac{[\mathcal{Y}(\theta_1^{*},\ldots,\theta_{i-1}^{*},
\theta_i^{*}+\Delta\theta_i^{*},
\theta_{i+1}^{*},\ldots,\theta_d^{*}) - 
\mathcal{Y}(\bm{\theta}^{*})]}{\Delta\theta_i^{*}} 
\label{eq:partial}
\ee

The total number of model realizations or function evaluations
needed to
compute $\mu_i$ in a Euclidian space $\mathbb{R}^d$ using $N$ samples is
therefore, $N\times(d+1)$. The derivative in Eq.~\ref{eq:partial} could be evaluated
analytically in case the functional dependence of $\mathcal{Y}$ and $\theta_i$
is known. Note that the perturbation $\Delta\theta_i^{*}$ is extremely small 
compared to the uncertainty in $\theta_i$ which leads to a significant improvement
over the `elementary effect' of a parameter, estimated in the Morris method~\cite{Morris:1991}
using large increments.  It is noted by many authors~[REFS], and also observed
in the numerical experiments in the present work, that a modest Monte Carlo
sample size is often sufficient for computing~\eqref{eq:mu} with reasonable
accuracy.

@@

\subsection{Polynomial Chaos Expansion}

As discussed earlier in Section~\ref{sec:intro}, the Polynomial Chaos Expansion (PCE)
is one of the most commonly used surrogates for representing the dependence of a random
observable ($\mathcal{Y}$) on independent uncertain or stochastic model parameters ($\bm{\theta}$).
Considering a joint probability distribution of the components of ($\bm{\theta}$) as $\mathbb{P}(\bm{\theta})$,
the following condition is imposed on $\mathcal{Y}$:

\be
\mathbb{E}[\mathcal{Y}^2] = \int_{\mathcal{D}_{\bm{\theta}}} \mathcal{Y}^2 \mathbb{P}(\bm{\theta}) 
d\bm{\theta} < \infty
\ee

\noindent where $\mathcal{D}_{\bm{\theta}}$ is the domain of the input parameter space. The PCE of
$\mathcal{Y}$ ($\mathcal{Y}^{\mbox{\tiny {PC}}}$) is a mean-square 
convergent truncated summation given as
follows~\cite{Xiu:2002,Ghanem:2003,Olivier:2010}:

\be
\mathcal{Y}^{\mbox{\tiny M}} \approx \mathcal{Y}^{\mbox{\tiny PC}} = 
\sum_{\alpha\in\mathcal{I}} c_{\alpha}\Psi_{\alpha}(\bm{\xi(\theta)}) 
\ee

\noindent where $\mathcal{Y}^{\mbox{\tiny M}}$ corresponds to realizations of the
observable using the original (typically physics-based) model. 
Note that the set of parameters, $\bm{\theta}$ are parameterized in terms of canonical random
variables, $\bm{\xi}$ in the above expansion. $\Psi_{\alpha}$'s are multivariate polynomial
basis, orthonormal with respect to the joint probability distribution of $\bm{\xi}$, $\mathcal{I}$
is the multi-index set which identifies the degree of individual polynomials in each term in the
multivariate basis.

@@

\be
\bm{c} = \mathbb{E}[\Psi_\alpha(\bm{\xi})\cdot\mathcal{Y}^{\mbox{\tiny{M}}}]
\approx
\sum_{i=1}^{N} \mathcal{Y}^{\mbox{\tiny{M}}}(\bm{\theta}^{(i)})w^{(i)}\Psi_\alpha(\bm{\xi}^{(i)})
\ee

@@

\be
\hat{\bm{c}} = \mbox{argmin}~\mathbb{E}\left[\left(\bm{c}^{T}\Psi(\bm{\bm{\xi(\theta)}}) -
\mathcal{Y}^{\mbox{\tiny{M}}}(\bm{\theta})\right)^{2}\right]  + \lambda\normone{\bm{c}}
\ee

>>>>

\subsection{Derivative-based global sensitivity analysis}

\alennote{1. Remove $G^M$; use $G$ and $G^{PC}$. 2. Start with $G$ in $\xi$ and
the, in one sentence connect to physical range. 3. Equation (9) should be moved up to 
2.1.}
Let $\GG$ be a mathematical model that is a function of $\Np$ uncertain 
parameters, $\theta_1, \theta_2, \ldots, \theta_\Np$. The goal of sensitivity analysis
is measuring the influence of each component of the parameter vector 
$\bm{\theta}$ on the model output. 
In the present work, we consider the case where the input parameters are statistically 
independent.

Derivative-based global sensitivity analysis (DGSA) is performed by 
computing derivative based global sensitivity measures (DGSMs)~\cite{Sobol:2009} 
for each uncertain parameter in the model. 
Specifically, we consider the following DGSMs, 
\be
\mu_i = 
\mathbb{E}\left[\left(\frac{\partial \GG(\bm{\bm{\theta}})}{\partial \theta_i}\right)^{2}\right], \quad i = 1, \ldots, \Np.
\label{eq:mu}
\ee
Here $\mathbb{E}$ denotes expectation over the uncertain parameters.
Notice that this formulation assumes that the function $\GG$ is differentiable
with respect to $\theta_i$, $i = 1, \ldots, \Np$, almost surely.

@@

\be
\frac{\partial \GG(\bm{\theta}^{\ast})}{\partial \theta_i} =
 \lim_{\Delta\theta_i^{*}\to 0}
\frac{[\GG(\theta_1^{*},\ldots,\theta_{i-1}^{*},
\theta_i^{*}+\Delta\theta_i^{*},
\theta_{i+1}^{*},\ldots,\theta_d^{*}) - 
\GG(\bm{\theta}^{*})]}{\Delta\theta_i^{*}} 
\label{eq:partial}
\ee

The total number of model realizations or function evaluations
needed to
compute $\mu_i$ in a Euclidian space $\mathbb{R}^d$ using $N$ samples is
therefore, $N\times(d+1)$. The derivative in Eq.~\ref{eq:partial} could be evaluated
analytically in case the functional dependence of $\GG$ and $\theta_i$
is known. Note that the perturbation $\Delta\theta_i^{*}$ is extremely small 
compared to the uncertainty in $\theta_i$ which leads to a significant improvement
over the `elementary effect' of a parameter, estimated in the Morris method~\cite{Morris:1991}
using large increments.  It is noted by many authors~[REFS], and also observed
in the numerical experiments in the present work, that a modest Monte Carlo
sample size is often sufficient for computing~\eqref{eq:mu} with reasonable
accuracy.

@@

\subsection{Polynomial chaos expansion}

%As discussed earlier in Section~\ref{sec:intro}, the 
%Polynomial chaos expansion
%(PCE) is a commonly used tool for surrogate modeling.
We consider models with $\Np$ random inputs, 
$\theta_1, \ldots, \theta_\Np$ that are modeled
as statistically independent random variables. The 
variables $\theta_i$ will take in physically meaningful
ranges; it is common to parameterize input uncertainties
with canonical random variables $\xi_1, \ldots, \xi_\Np$,
which can be then shifted and scaled to obtain the corresponding $\theta_i's$.
Typical choices for distribution of $\xi_i$ include standard normal 
and uniform distribution on the interval $[-1, 1]$.
Let 
\[
   \bm{f}(\bm{x}) = \prod_{i=1}^\Np f_i(x_i), \quad \bm{x} \in \mathbb{R}^\Np
\]
where $f_i$ are probability density functions of $\xi_i$, $i = 1, \ldots, \Np$.
We consider a square integrable random variable $\GG:\R^\Np \to \R$; 
that is,
\[
\int_{\mathcal{D}} \GG(\bm{\xi})^2 \, \bm{f}(\bm{\xi})d\bm{\xi} < \infty,
\]
where $\mathcal{D}$ is the support of the distribution law of the random vector
$\bm{\xi}$. In applications, 
the random variable $\GG$ is the mathematical model of interest.

% for representing the
%dependence of a random observable ($\GG$) on independent uncertain or
%stochastic model parameters ($\bm{\theta}$).  Considering a joint probability
%distribution of the components of ($\bm{\theta}$) as
%$\mathbb{P}(\bm{\theta})$, the following condition is imposed on
%$\GG$:

\be
\mathbb{E}[\GG^2] = \int_{\mathcal{D}_{\bm{\theta}}} \GG^2 \mathbb{P}(\bm{\theta}) 
d\bm{\theta} < \infty
\ee

\noindent where $\mathcal{D}_{\bm{\theta}}$ is the domain of the input parameter space. The PCE of
$\GG$ ($\GG^{\mbox{\tiny {PC}}}$) is a mean-square 
convergent truncated summation given as
follows~\cite{Xiu:2002,Ghanem:2003,Olivier:2010}:

\be
\GG^{\mbox{\tiny M}} \approx \GG^{\mbox{\tiny PC}} = 
\sum_{\alpha\in\mathcal{I}} c_{\alpha}\Psi_{\alpha}(\bm{\xi(\theta)}) 
\ee

\noindent where $\GG^{\mbox{\tiny M}}$ corresponds to realizations of the
observable using the original (typically physics-based) model. 
Note that the set of parameters, $\bm{\theta}$ are parameterized in terms of canonical random
variables, $\bm{\xi}$ in the above expansion. $\Psi_{\alpha}$'s are multivariate polynomial
basis, orthonormal with respect to the joint probability distribution of $\bm{\xi}$, $\mathcal{I}$
is the multi-index set which identifies the degree of individual polynomials in each term in the
multivariate basis.

@@

\be
\bm{c} = \mathbb{E}[\Psi_\alpha(\bm{\xi})\cdot\GG^{\mbox{\tiny{M}}}]
\approx
\sum_{i=1}^{N} \GG^{\mbox{\tiny{M}}}(\bm{\theta}^{(i)})w^{(i)}\Psi_\alpha(\bm{\xi}^{(i)})
\ee

@@

\be
\hat{\bm{c}} = \mbox{argmin}~\mathbb{E}\left[\left(\bm{c}^{T}\Psi(\bm{\bm{\xi(\theta)}}) -
\GG^{\mbox{\tiny{M}}}(\bm{\theta})\right)^{2}\right]  + \lambda\normone{\bm{c}}
\ee
